---
title: "PA1_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Reproducible Research Peer-graded Assignment: Course Project 1 (week 2) 

### Loading and preprocessing the data
1.Load the data (i.e. read.csv())
```
mainsource <- read.csv("./activity.csv", sep = ",", header = TRUE, na.strings = "NA")
```
2.Process/transform the data (if necessary) into a format suitable for your analysis
```
mainsource$date<-as.Date(mainsource$date,"%Y-%m-%d")
```

###What is mean total number of steps taken per day?
1.Calculate the total number of steps taken per day
```
totalnosteps <-aggregate(mainsource$steps, list(mainsource$date), sum, na.rm = TRUE)
head(totalnosteps,10)
```
2.Make a histogram of the total number of steps taken each day
```
hist(as.numeric(totalnosteps$steps),
                  breaks = 10,
                  col = "blue",
                  xlab = "No. of Steps Taken",
                  main="Histogram of the Total Number of Steps Taken Each Day")
```
```{r, echo=FALSE}
setwd("d:\\coursera")
mainsource <- read.csv("./activity.csv", sep = ",", header = TRUE, na.strings = "NA")
mainsource$date<-as.Date(mainsource$date,"%Y-%m-%d")
totalnosteps <- aggregate(mainsource$steps, list(mainsource$date),sum)
colnames(totalnosteps) <- c("date","steps")
hist(as.numeric(totalnosteps$steps),
                  breaks = 20,
                  col = "blue",
                  xlab = "No. of Steps Taken",
                  main="Histogram of the Total Number of Steps Taken Each Day")
```

3.Calculate and report the mean and median of the total number of steps taken per day
```
Meantotalnosteps <- mean(totalnosteps$steps, na.rm = TRUE)
```{r}
mean(totalnosteps$steps, na.rm = TRUE)
```
```
Mediantotalnosteps <- median(totalnosteps$steps, na.rm = TRUE)
```{r}
median(totalnosteps$steps, na.rm = TRUE)
```

###What is the average daily activity pattern?
1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```
stepsinterval <- aggregate(steps ~ interval, data = mainsource, mean)
plot(stepsinterval$interval, stepsinterval$steps, 
                  col = "blue", 
                  type = "l", 
                  xlab = "5 Minute Interval", 
                  ylab = "Number of Steps",
                  main = "Average Number of Steps Taken")
```
```{r, echo = FALSE}
stepsinterval <- aggregate(steps ~ interval, data = mainsource, mean)
plot(stepsinterval$interval, stepsinterval$steps, 
                  col = "blue", 
                  type = "l", 
                  xlab = "5 Minute Interval", 
                  ylab = "Number of Steps",
                  main = "Average Number of Steps Taken")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```
maxsteps <- stepsinterval[which.max(stepsinterval$steps), ]
```{r}
stepsinterval[which.max(stepsinterval$steps), ]
```
It is observed that the maximum number of steps is 206 which occurs at interval 835.

###Imputing missing values
1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(mainsource$steps))
sum(is.na(mainsource$date))
sum(is.na(mainsource$interval))
```
Total number of missing values in the dataset is 2304.

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```
missingvalue <- which(is.na(mainsource$steps))
fillNA <- mainsource
fillNA[missingvalue, ]$steps<-unlist(lapply(missingvalue, FUN=function(missingvalue){
                stepsinterval[mainsource[missingvalue,]$interval==stepsinterval$interval,]$steps
                }))
```
3.Creating a new dataset that is equal to the original dataset but with the missing data filled in.
```
str(fillNA)
```
```{r, echo = FALSE}
missingvalue <- which(is.na(mainsource$steps))
fillNA <- mainsource
fillNA[missingvalue, ]$steps<-unlist(lapply(missingvalue, FUN=function(missingvalue){
                stepsinterval[mainsource[missingvalue,]$interval==stepsinterval$interval,]$steps
                }))
str(fillNA)
```
4.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```
totalnostepsfillNA <- aggregate(steps ~ date, data = fillNA, sum)
hist(as.numeric(totalnostepsfillNA$steps),
                  breaks = 20,
                  col = "blue",
                  xlab = "No. of Steps Taken",
                  main="Histogram of the Total Number of Steps Taken Each Day")
                  
```
```{r, echo = FALSE }
missingvalue <- which(is.na(mainsource$steps))
fillNA <- mainsource
fillNA[missingvalue, ]$steps<-unlist(lapply(missingvalue, FUN=function(missingvalue){
                stepsinterval[mainsource[missingvalue,]$interval==stepsinterval$interval,]$steps
                }))
totalnostepsfillNA <- aggregate(steps ~ date, data = fillNA, sum)
hist(as.numeric(totalnostepsfillNA$steps),
                  breaks = 20,
                  col = "blue",
                  xlab = "No. of Steps Taken",
                  main="Histogram of the Total Number of Steps Taken Each Day")
                  
```
```{r}
mean(totalnostepsfillNA$steps, na.rm = TRUE)
median(totalnostepsfillNA$steps, na.rm = TRUE)
```
Do these values differ from the estimates from the first part of the assignment? Yes, the values differ from the first part because for this part, the mean and median are both equal at 10766.19

What is the impact of imputing missing data on the estimates of the total daily number of steps? The impact of imputting missing data changes the median.

###Are there differences in activity patterns between weekdays and weekends?
1.Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```
fillNA$day<-as.factor(weekdays(fillNA$date))
fillNA$indicator<-ifelse(fillNA$day %in% c("Saturday","Sunday"), "weekend", "weekday")
head(fillNA)
```
2.Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend fdays (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```
#the average number of steps taken for weekdays 
weekdaysfillNA <- fillNA[fillNA$indicator == weekday",]
stepsperweekday <- aggregate(weekdaysfillNA$steps, list(weekdaysfillNA$interval), FUN=mean)
stepsperweekday$indicator <- "weekday"

#the average number of steps taken for weekends
weekendfillNA <- fillNA[fillNA$indicator == "weekend",]
stepsperweekend <- aggregate(weekendfillNA$steps, list(weekendfillNA$interval), FUN=mean)
stepsperweekend$indicator <- "weekend"

newfillNA <- rbind(stepsperweekday,stepsperweekend)
newfillNA$indicator <- as.factor(newfillNA$indicator)
colnames(newfillNA) <- c("interval","steps","indicator")
library(lattice)
xyplot(steps ~ interval  | indicator, data = newfillNA, layout = c(1,2), type ="l", ylab="Number of Steps")
```
```{r, echo = FALSE }
fillNA$day<-as.factor(weekdays(fillNA$date))
fillNA$indicator<-ifelse(fillNA$day %in% c("Saturday","Sunday"), "weekend", "weekday")

weekdaysfillNA <- fillNA[fillNA$indicator == "weekday",]
stepsperweekday <- aggregate(weekdaysfillNA$steps, list(weekdaysfillNA$interval), FUN=mean)
stepsperweekday$indicator <- "weekday"

weekendfillNA <- fillNA[fillNA$indicator == "weekend",]
stepsperweekend <- aggregate(weekendfillNA$steps, list(weekendfillNA$interval), FUN=mean)
stepsperweekend$indicator <- "weekend"

newfillNA <- rbind(stepsperweekday,stepsperweekend)
newfillNA$indicator <- as.factor(newfillNA$indicator)
colnames(newfillNA) <- c("interval","steps","indicator")
library(lattice)
xyplot(steps ~ interval  | indicator, data = newfillNA, layout = c(1,2), type ="l", ylab="Number of Steps")
```

Yes, there are differences in activity patterns between weekdays and weekends. As observed, weekend at most times exceeded the 100 steps mark but never reached the 200 steps mark. Weekday was able to reach the 200 steps mark at some point, but it remains below the 100 steps mark throughout the duration of the interval. 